\bf{Inequality in non-cognitive skills: \\ a cross-cohort comparison}

### 2018-05-22 codes/R/slides/cohorts_tapes.Rnw
slide update

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_scores.R"))
read_chunk(here("codes/R/4_scores_aux.R"))
read_chunk(here("codes/R/5_graphs.R"))
read_chunk(here("codes/R/6_selectvars.R"))
read_chunk(here("codes/R/8_regs.R"))
read_chunk(here("codes/R/getpars.R"))

@@

\bi
\item Non-cognitive (socio-emotional, behavioural) skills developed during childhood are important for lifetime outcomes
\item Socioeconomic gradients in these skills are persistent 
\pause 
\item \textbf{Q1}: Can we derive \textbf{comparable measures} across populations?
    \bi
    \item Usually, behavioural constructs are measured via scales
    \item But does the same scale measure the same constructs in the same way:
      \bi 
      \item across different populations?
      \item in the same population, at different points in time?
      \ei
    \item And what if only different scales are available?
    \ei
\pause
\item \textbf{Q2}: Can we use these measures to document \textbf{changes in socioeconomic inequality} in skills across populations?
\ei

@@

\begin{frame}
\frametitle{This study}

\bi
\item Focus on behavioural problems in two British cohorts
  \bi \item British Cohort Study (BCS, b. 1970)
  \item Millennium Cohort Study (MCS, b. 2000/1) \ei
\ei
\vspace{6pt} \pause
\bnm
\item Construct a common 11-item scale of non-cognitive skills, starting from different instruments
\item Investigate its psychometric properties
\item Use scores from the new scale to document changes in socioeconomic gradients between cohorts 30 years apart
\enm
\vspace{6pt} \pause
\bi
\item \textbf{Key findings}:
  \bi
  \item Inequality in non-cognitive skills has increased in 1975-2005
  \item Sharpest change for boys rather than girls
  \item Mostly driven by improvements at the top of the distribution
  \ei
\ei

@@

\subsection{Data}

@@

\begin{column}{0.6\textwidth}
\bi
\item Data from 5 year follow-up (England only), maternal questionnaire
\item Rutter A scale \citep{Rutter1970}
  \bi
  \item 19 items (behaviours)
  \item Options: \textit{Doesn't apply, Applies somewhat, Certainly applies}
  \item Usually conceptualised as measuring two dimensions
  \bi \item \emph{externalising} behaviour problems (anti-social) \item \emph{internalising} behaviour problems (neurotic) \ei
  \ei
  \item Additional questions from maternal module
\ei
\end{column}

@@

\bi
\item Data from 5 year follow-up (England only), maternal questionnaire
  \bi \item Complex survey design: rebalance sample to make representative \ei
\item Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994})
  \bi
  \item ``Updated" and expanded version of the Rutter scale 
  \item 25 items (behaviours)
  \item Options: \textit{Not true, Somewhat true, Certainly true}
  \item Designed with 5 subscales
  \item Broader subscales are sometimes justifiable \citep{Goodman2010}
  \ei
\ei

@@

@@

@@

table <- meantab[[2]] # females (but just need item names)
table <- table[,!names(table) %in% c("cats", "fac", "bcs.num", "mcs.num")] # drop if unneeded

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% EXPLORATORY FACTOR ANALYSIS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{frame}
\frametitle{EFA}\label{efa}
\framesubtitle{Factor structure}

\bi
\item Different methods for assessing number of factors give varied results \hyperlink{numfac}{\beamerbutton{Det.}}
\item We choose two factors to maximise interpretability
\item In both cohorts, EFA delivers a clear-cut split between two factors
\item Factors can be coherently interpreted as internalising and internalising skills
\ei 
\vspace{12pt}
\centering
\begin{tabular}{C{.05\textwidth}L{.3\textwidth}|C{.05\textwidth}L{.3\textwidth}}
  & \textbf{Externalising}  &    & \textbf{Internalising} \\
1 & Restless                & 7  & Worried                \\
2 & Squirmy/fidgety         & 8  & Afraid of new things   \\
3 & Fights/bullies children & 9  & Solitary               \\
4 & Cannot settle           & 10 & Unhappy/tearful        \\
5 & Temper tantrums         & 11 & Head/stomachaches \\
6 & Disobedient             &    &                       
\end{tabular}

@@

@@

\bi
\item Two constructs isolated in the new scale, separately by cohort
\item But are externalising and internalising problems measured in a comparable way \emph{across cohorts}?
\vspace{12pt}\pause
\item \textbf{Measurement Invariance (MI)} analysis
  \bi
  \item Formal test of the invariance hypothesis
  \item Set up categorical Item Factor Analysis (IFA) model \hyperlink{miapp1}{\beamerbutton{Det.}}
  \item Pool the two cohorts together, and estimate models of different invariance level -- following \cite{Wu2016}:\hyperlink{miapp2}{\beamerbutton{Det.}}
    \bnm
    \item Start from minimally restrictive model, where parameters are free to vary
    \item Restrict threshold and loading parameters to be equal across cohorts
    \item Restrict intercept parameters as well
    \enm
  \item Test relative fit of restricted models compared to initial model
  \ei 
\ei

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% MEASUREMENT INVARIANCE RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\bi
\item Restricting thresholds and loadings provides a good fit \hyperlink{fit}{\beamerbutton{Fit}}
  \bi \item We can compare the scale (variance) of the latent factors \ei
\item Models with restricted intercepts do \emph{not} pass MI test
  \bi
  \item Unclear whether mean comparisons can be done
  \item \cite{Wu2016} vs \cite{Millsap2004}
  \ei
% \item Might \emph{age differences} in the sample be playing a role?
%   \bi
%   \item BCS and MCS samples differ due to survey design \hyperlink{agehist}{\beamerbutton{Graph}}
%   \item But restricting the age range to maximise overlap does not give significantly different results \hyperlink{fit4}{\beamerbutton{Det.}}
%   \ei
  \pause
\item \textbf{Q1}: Can we derive comparable measures of noncognitive skills across cohorts?
\item Yes, we can use the model with non-invariant intercepts to score the latent factors (externalising and internalising skills)
\ei

@@

\begin{frame}
\frametitle{Inequality}\label{ineq_sc}
\framesubtitle{Age mother left education}
\vspace{-.2cm}
\begin{figure}
<<FACINEQ_YS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=7.5, fig.width=10, out.width='.9\\linewidth'>>=
# plot
fi <- fi_ys # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}\label{ineq_sc}
\framesubtitle{Maternal education and employment}
\vspace{-.2cm}
\begin{figure}
<<FACINEQ_MECS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=8.5, fig.width=13.5, out.width='\\linewidth'>>=
# plot
fi <- fi_mecs # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{What explains behavioural skills?}
\framesubtitle{Males}
\begin{table}
\centering
\resizebox{!}{4cm}{
<<REGS_M, results='asis', dependson="REGRESSIONS">>=

@@

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = robust_se, override.pvalues = robust_p,
               single.row = TRUE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = "region|Inter",
               groups = grouplabs,
               custom.coef.names = varlabs,
               custom.model.names = c("EXT BCS", "EXT MCS", "INT BCS", "INT MCS"),
               table = FALSE
               )

@@

\begin{frame}
\frametitle{What explains behavioural skills?}
\framesubtitle{Females}
\begin{table}
\centering
\resizebox{!}{4cm}{
<<REGS_F, results='asis', dependson="REGRESSIONS">>=

@@

\begin{frame}
\frametitle{Decomposition (EXT)}
\framesubtitle{Explaining the education gradient}
\centering
\scriptsize
\resizebox{\textwidth}{!}{
\input{../../../tables/med_gelb_ext}
}
\end{frame}

\begin{frame}
\frametitle{Decomposition (INT)}
\framesubtitle{Explaining the education gradient}
\centering
\scriptsize
\resizebox{\textwidth}{!}{
\input{../../../tables/med_gelb_int}
}
\end{frame}

@@

@@

# collapse to matrix (first 4 elements of bcs_outctab are adolescence)
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c(1:4)],matrix,ncol=11,byrow=FALSE))
table <- bcs_outctab_mat

print(xtable(table,
                    align="R{0cm}L{.22\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}",
                    digits = 3
                    ),
             size="\\scriptsize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            " & \\multicolumn{5}{c}{\\textbf{Males}} & \\multicolumn{5}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-6}  \\cmidrule(lr){7-11}
& Mean & \\multicolumn{4}{c}{Coefficients} & Mean & \\multicolumn{4}{c}{Coefficients} \\\\ \\cmidrule(lr){3-6}  \\cmidrule(lr){8-11}"
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
@
}
\end{frame}

@@

print(xtable(table,
                    align="R{0cm}L{.22\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}",
                    digits = 3
                    ),
             size="\\scriptsize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            " & \\multicolumn{5}{c}{\\textbf{Males}} & \\multicolumn{5}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-6}  \\cmidrule(lr){7-11}
& Mean & \\multicolumn{4}{c}{Coefficients} & Mean & \\multicolumn{4}{c}{Coefficients} \\\\ \\cmidrule(lr){3-6}  \\cmidrule(lr){8-11}"
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
@
}
\end{frame}

print(xtable(table,
                    align="R{0cm}L{.22\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.05\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}C{.075\\textwidth}",
                    digits = 3
                    ),
             size="\\scriptsize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            " & \\multicolumn{5}{c}{\\textbf{Males}} & \\multicolumn{5}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-6}  \\cmidrule(lr){7-11}
& Mean & \\multicolumn{4}{c}{Coefficients} & Mean & \\multicolumn{4}{c}{Coefficients} \\\\ \\cmidrule(lr){3-6}  \\cmidrule(lr){8-11}"
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
@
}
\end{frame}

@@

@@

<<BCSEFA>>=
@

<<BCSEFARES, dependson='BCSEFA'>>=
@

<<MCSEFA>>=
@

<<MCSEFARES, dependson='MCSEFA'>>=
@

\begin{frame}[noframenumbering]
\frametitle{EFA}\label{numfac}
\framesubtitle{Number of factors \hyperlink{efa}{\beamerreturnbutton{Back}}}
\begin{center}
<<EFATAB, dependson='MCSEFA', results='asis'>>=
# combine results
efatab <- cbind(efatab.bcs,efatab.mcs)
colnames(efatab) <- c("BCS All", "BCS M", "BCS F", "MCS All", "MCS M", "MCS F")

print(xtable(efatab,
                    align="L{.25\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 0
                    ),
             size="\\small",
             include.rownames=T,
             include.colnames=T,
             booktabs = T,
             sanitize.text.function = function(x){x},
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)
@
\end{center}
\end{frame}

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp1}
\framesubtitle{IFA model}
Categorical Item Factor Analysis (IFA) model \hyperlink{mimain}{\beamerreturnbutton{Back}}

\bi
\item Threshold model \citep{Muthen1984}
\item Latent propensity to endorse each item $i$ for child $j$ in cohort $c$
\begin{align}\label{eq:propens}
\begin{split}
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{EXT}_{jc} + u_{ijc} \qquad \text{for} \; i=1,\dots, 6 \\
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{INT}_{jc} + u_{ijc} \qquad \text{for} \; i=7,\dots, 11
\end{split}
\end{align}

\begin{frame}[noframenumbering]
\frametitle{MI}
\framesubtitle{Key Parameters for Invariance}

\bi
\item Loadings $\lambda_{ci}$
\item Thresholds $\tau_{1,ci}$, $\tau_{2,ci}$
\item Intercepts $\nu_{ci}$
\item Unique variances $\Var(u_{ci})$
\item Latent mean vector $\bm{\kappa}_c$
\item Latent covariance matrix $\bm{\Phi}_c$
\ei

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp2}
\framesubtitle{Invariance steps \hyperlink{mimain}{\beamerreturnbutton{Back}}}
\Fontvi
\bi
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across cohorts)
  \item $\text{diag}(\bm{\Phi}) = \bm{I}, \quad \bm{\kappa}=\bm{0}, \quad \bm{\nu}=\bm{0}, \quad \text{and} \; \text{diag}(\bm{\Psi})=\bm{I}.$
  \ei

>>>>

<<READ, cache=FALSE >>=
library(here)
read_chunk(here("codes/R/1_load.R"))
read_chunk(here("codes/R/2_efa.R"))
read_chunk(here("codes/R/3_inv_main.R"))
read_chunk(here("codes/R/3_inv_aux.R"))
read_chunk(here("codes/R/3_inv_fit.R"))
read_chunk(here("codes/R/4_scores.R"))
read_chunk(here("codes/R/4_scores_aux.R"))
read_chunk(here("codes/R/5_graphs.R"))
read_chunk(here("codes/R/6_selectvars.R"))
read_chunk(here("codes/R/8_regs_dep.R"))
read_chunk(here("codes/R/8_regs_outc.R"))
read_chunk(here("codes/R/getpars.R"))

@@

\bi
\item Socio-emotional skills developed during childhood are important for lifetime outcomes
\item Socioeconomic gradients in these skills are persistent 
\item {\color{red} IMPROVE INTRO...}
\ei

@@

\begin{frame}
\frametitle{This study}
\framesubtitle{Research questions}
\bi
\item Usually, socioemotional constructs are measured using validated scales
    \bi \item e.g. Rutter, BSAG, SDQ \ei
\pause
~\\    
\item \textbf{Q1}: Can we derive \textbf{comparable measures} of socioemotional skills across populations?
    \bi
    \item Does the same instrument measure the same constructs in the same way:
      \bi 
      \item across different populations?
      \item in the same population, at different points in time?
      \ei
    \item And what if only different instruments are available?
    \ei
\pause
~\\    
\item \textbf{Q2}: \textbf{How many latent dimensions} are underlying these instruments?
    \bi
    \item Can they be simply summarised into a single score?
    \ei
\ei
\end{frame}

\begin{frame}
\frametitle{This study}
\framesubtitle{Overview}

\bi
\item Focus on socioemotional skills in two British cohorts
  \bi \item British Cohort Study (BCS, b. 1970)
  \item Millennium Cohort Study (MCS, b. 2000/1) \ei
\ei
~\\ \pause
\textbf{Contribution}
\bi
\item Construct a common 11-item scale, starting from different survey instruments
\item Investigate its psychometric properties, and find
  \bnm 
  \item \textbf{\textit{Comparability}} in terms of scale (but not location)
  \item \textbf{\textit{Bidimensionality}} -- Externalising and Internalising
  \enm
\ei
~\\ \pause

\textbf{Key findings}
  \bi
  \item Inequality in non-cognitive skills has increased in 1975-2005
  \item Sharpest change for boys rather than girls
  \item {\color{red} OTHER RESULTS}
  \ei

@@

\section{Data}

\begin{frame}
\frametitle{Data}
\framesubtitle{The UK birth cohort studies}

\bi
\item Four main cohorts:
  \bi
  \item \textit{National Survey of Health and Development} (NSHD) -- b. March 1946
  \item \textit{National Child Development Study} (NCDS) -- b. March 1958
  \item \textit{British Cohort Study} (BCS) -- b. April 1970
  \item \textit{Millennium Cohort Study} (MCS) -- b. 2000-2001
  \ei
~\\
\item Follow 15,000+ individuals from birth onwards
\item Include extensive data on
  \bi
  \item Perinatal period (pregnancy, birth outcomes)
  \item Skill formation -- cognitive, socioemotional
  \item Health and health behaviours
  \item Family background and composition
  \item Schooling and labour market outcomes
  \ei
~\\
\pause
\item Not always consistent
  \bi
  \item Different design, timing, questions, survey instruments...
  \item Harmonisation efforts (CLOSER)
  \ei
\ei

@@

\begin{column}{0.6\textwidth}
\bi
\item Data from 5 year follow-up
\item Sample restricted to England
\item Socioemotional skills: \textbf{Rutter A scale} \citep{Rutter1970}
  \bi
  \item 19 items (descriptions of behaviours) from maternal questionnaire
  \item Options: \textit{Doesn't apply, Applies somewhat, Certainly applies}
  \ei
  \item Additional questions from maternal module
\ei
\end{column}

@@

\bi
\item Data from 5 year follow-up
\item Sample restricted to England
\item Complex survey design: rebalance sample to make representative
\item Socioemotional skills: \textbf{Strengths and Difficulties Questionnaire} (SDQ, \citealp{Goodman1994})
  \bi
  \item ``Updated" and expanded version of the Rutter scale 
  \item 25 items (descriptions of behaviours)
  \item Options: \textit{Not true, Somewhat true, Certainly true}
  \ei
\ei

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% HARMONISATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{frame}
\frametitle{Harmonisation}
\framesubtitle{Choice of common items}

\bi
\item Rutter and SDQ are different instruments
\item Focus on subset of items that have similar wording between the two instruments
\item New 11-item subscale
\ei

@@

table <- meantab[[2]] # females
table <- table[,!names(table) %in% c("cats", "fac", "bcs.num", "mcs.num")] # drop if unneeded

@@

@@

\begin{frame}
\frametitle{Factor structure}
\framesubtitle{Existing evidence}

@@

\begin{frame}
\frametitle{EFA}\label{efa}
\framesubtitle{Dimensionality}

\bi
\item Different methods for assessing number of factors give varied results \hyperlink{numfac}{\beamerbutton{Det.}}
\item We choose two factors to maximise interpretability
\item In both cohorts, EFA delivers a clear-cut split between two factors
\item Factors can be coherently interpreted as internalising and internalising skills
\ei 
\vspace{12pt}
\centering
\begin{tabular}{C{.05\textwidth}L{.3\textwidth}|C{.05\textwidth}L{.3\textwidth}}
  & \textbf{{\color{OliveGreen}Externalising}}  &    & \textbf{{\color{Orange}Internalising}}  \\
1 & Restless                & 7  & Worried                \\
2 & Squirmy/fidgety         & 8  & Afraid of new things   \\
3 & Fights/bullies children & 9  & Solitary               \\
4 & Cannot settle           & 10 & Unhappy/tearful        \\
5 & Temper tantrums         & 11 & Head/stomachaches \\
6 & Disobedient             &    &                       
\end{tabular}

\begin{frame}
\frametitle{EFA}\label{meantab_fcc2}
\framesubtitle{Choice of common items (females) \hyperlink{meantab_fcc3}{\beamerbutton{3 factors}}}
\centering
\resizebox{!}{4cm}{
<<MEANTAB_FCC, dependson="MEANTABLE", results='asis'>>=
require(xtable)

table <- meantab_cc2[[2]] # females
table <- table[,!names(table) %in% c("cats", "fac", "bcs.num", "mcs.num")] # drop if unneeded

@@

\bi
\item Two constructs isolated in the new scale, separately by cohort
\item But are externalising and internalising problems measured in a comparable way \emph{across cohorts}?
\vspace{12pt}\pause
\item \textbf{Measurement Invariance (MI)} analysis
  \bi
  \item Formal test of the invariance hypothesis
  \item Set up categorical Item Factor Analysis (IFA) model
  \item Pool the two cohorts together, and estimate models of different invariance level -- following \cite{Wu2016}
  \item Test relative fit of restricted models compared to initial model
  \ei 
\ei

@@

\begin{frame}
\frametitle{Measurement Invariance}\label{midet}
\framesubtitle{IFA model}
Categorical Item Factor Analysis (IFA) model

\bi
\item Threshold model \citep{Muthen1984}
\item Latent propensity to endorse each item $i$ for child $j$ in group $c$
\begin{align}\label{eq:propens}
\begin{split}
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{EXT}_{jc} + u_{ijc} \qquad \text{for} \; i=1,\dots, 6 \\
X^*_{ijc} &=\nu_{ci} + \lambda_{ci}\theta^{INT}_{jc} + u_{ijc} \qquad \text{for} \; i=7,\dots, 11
\end{split}
\end{align}

\begin{frame}
\frametitle{MI}\label{misteps}
\framesubtitle{Invariance steps \hyperlink{miapp1}{\beamerbutton{Details}}}
\bnm
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across groups)
  \item Normalised on latent means $\bm{\kappa}$ and variances $\bm{\Phi}$
  \ei
\item \textbf{Threshold and Loading invariance} model
  \bi
  \item Restrict $\bm{\tau}_{1}$, $\bm{\tau}_{2}$, and $\bm{\Lambda}$ to be equal across groups
  \item Allows comparison of latent factor variances
  \ei
\item \textbf{Threshold, Loading, and Intercept invariance} model
  \bi
  \item Restrict $\bm{\tau}_{1}$, $\bm{\tau}_{2}$, $\bm{\Lambda}$, and $\bm{\nu}$ to be equal across groups
  \item Allows comparison of latent factor variances \emph{and} means
  \ei
\enm
\end{frame}

@@

@@

\bi
\item We use RMSE to compare fit between models of different invariance level
\item Restricting thresholds and loadings provides a good fit \hyperlink{fit}{\beamerbutton{Fit details}}
  \bi \item We can compare the scale (variance) of the latent factors \ei
\item Models with restricted intercepts do \emph{not} pass MI test
  \bi
  \item Unclear whether mean comparisons can be done
  \item \cite{Wu2016} vs \cite{Millsap2004}
  \ei
\ei
\begin{center}
<<FA_FIT_RES, dependson="FA_FIT", results='asis'>>=
require(xtable)

table <- rbind(afitab[[1]])
table <- as.matrix(table[,c("mnames","npar","rmsea")])

for (c in 1:ncol(table)) {
  table[2,c] <- paste0("{\\color{green}", table[2,c], "}")
  table[3,c] <- paste0("{\\color{red}", table[3,c], "}")
}

print(xtable(table,
                    align="L{0cm}L{.3\\textwidth}C{.1\\textwidth}C{.15\\textwidth}",
                    digits = 4
                    ),
             size="\\footnotesize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            "Model & Num. params & RMSEA \\\\"
                                    )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "5\\textwidth",
             floating = FALSE
)

@
\end{center}
\end{frame}

\begin{frame}
\frametitle{Measurement Invariance}
\framesubtitle{Results \hyperlink{measpars}{\beamerbutton{Measurement}}\hyperlink{lvpars}{\beamerbutton{Latent variable}}}
\bi
\item \textbf{Q1}: Can we derive comparable measures of noncognitive skills across cohorts?
\item Yes, we can use the model with non-invariant intercepts to score the latent factors (externalising and internalising skills)
\item All scores are coded so that \emph{higher is better}
\ei
\end{frame}

@@

\begin{frame}
\frametitle{Inequality}\label{ineq_ys}
\framesubtitle{Age mother left education}
\vspace{-.2cm}
\begin{figure}
<<FACINEQ_YS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=7.5, fig.width=10, out.width='.9\\linewidth'>>=
# plot
fi <- fi_ys # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}\label{ineq_mecs}
\framesubtitle{Maternal education and employment}
\vspace{-.2cm}
\begin{figure}
<<FACINEQ_MECS, dependson="FA_SCORES_BIND", dev='tikz', fig.height=7.5, fig.width=11, out.width='\\linewidth'>>=
# plot
fi <- fi_mecs # select result
gridExtra::grid.arrange(fi$ineqlist[[1]],fi$ineqlist[[2]], 
                        fi$ineqlist[[3]],fi$ineqlist[[4]],
                        fi$histlist[[1]],fi$histlist[[2]], fi$legend,
                        ncol=2, nrow=4, heights = c(4,4,2,0.5))
@
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{What explains behavioural skills?}
\framesubtitle{Males}
\begin{table}
\centering
\resizebox{\textwidth}{!}{
<<REGS_M, results='asis', dependson="REGRESSIONS">>=

@@

texreg::texreg(regs,
               stars = c(0.1, 0.05, 0.01), digits = 3,
               override.se = robust_se, override.pvalues = robust_p,
               single.row = TRUE, scriptsize = TRUE,
               caption = "", label = "",
               omit.coef = "region|Inter",
               groups = grouplabs,
               custom.coef.names = varlabs,
               custom.model.names = c("EXT BCS", "EXT MCS", "INT BCS", "INT MCS"),
               include.rmse = FALSE,
               table = FALSE
               )

@@

\begin{frame}
\frametitle{What explains behavioural skills?}
\framesubtitle{Females}
\begin{table}
\centering
\resizebox{\textwidth}{!}{
<<REGS_F, results='asis', dependson="REGRESSIONS">>=

@@

\begin{frame}
\frametitle{Decomposition (EXT)}\label{gelb_ext}
\framesubtitle{Explaining the education gradient \hyperlink{gelbsc_ext}{\beamerbutton{Social class}}}
\centering
\scriptsize
\resizebox{\textwidth}{!}{
\input{../../../tables/med_gelb_ext}
}
\end{frame}

\begin{frame}
\frametitle{Decomposition (INT)}\label{gelb_int}
\framesubtitle{Explaining the education gradient \hyperlink{gelbsc_int}{\beamerbutton{Social class}}}
\centering
\scriptsize
\resizebox{\textwidth}{!}{
\input{../../../tables/med_gelb_int}
}
\end{frame}

@@

<<REGS_OUTC_TABPR, dependson="REGS_OUTC">>=
# function to display outcome regression tables
printoutctab <- function(tab) {
print(xtable(tab,
                    align="R{0cm}L{.2\\textwidth}C{.05\\textwidth}C{.09\\textwidth}C{.09\\textwidth}C{.09\\textwidth}C{.09\\textwidth}C{.05\\textwidth}C{.09\\textwidth}C{.09\\textwidth}C{.09\\textwidth}C{.09\\textwidth}",
                    digits = 3
                    ),
             size="\\scriptsize",
             include.rownames=F,
             include.colnames=F,
             booktabs = T,
             add.to.row = list(pos = c(list(0)),
                              command = c(
                              # header
            " & \\multicolumn{5}{c}{\\textbf{Males}} & \\multicolumn{5}{c}{\\textbf{Females}} \\\\ \\cmidrule(lr){2-6}  \\cmidrule(lr){7-11}
& Mean & \\multicolumn{4}{c}{Coefficients} & Mean & \\multicolumn{4}{c}{Coefficients} \\\\ \\cmidrule(lr){3-6}  \\cmidrule(lr){8-11}"
            )),
             sanitize.text.function = function(x){x},
             sanitize.rownames.function = function(x){""}, # to not print row numbers
             tabular.environment = "tabular", width = "\\textwidth",
             floating = FALSE
)
}
@

# NAs in columns and rows with COG
table[,c(6,11)] <- NA
table[seq(4,16,4),] <- NA

printoutctab(table)

\begin{frame}[noframenumbering]
\frametitle{Outcomes}
\framesubtitle{Adolescence (MCS)}
\centering
\resizebox{\textwidth}{!}{
<<REGS_OUTC_MCS_COG, results='asis', dependson="REGRESSIONS">>=

printoutctab(table)

@@

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c("smktry16", "alcoh16", "bmi16")],matrix,ncol=11,byrow=FALSE))
table <- bcs_outctab_mat

# NAs in columns and rows with COG
table[,c(6,11)] <- NA
table[seq(4,12,4),] <- NA

printoutctab(table)

\begin{frame}[noframenumbering]
\frametitle{Outcomes}
\framesubtitle{Adolescence (BCS)}
\centering
\resizebox{\textwidth}{!}{
<<REGS_OUTC_BCS1_COG, results='asis', dependson="REGRESSIONS">>=

# collapse to matrix
bcs_outctab_mat <- do.call(rbind,lapply(bcs_outctab[c("smktry16", "alcoh16", "bmi16")],matrix,ncol=11,byrow=FALSE))
table <- bcs_outctab_mat

printoutctab(table)

@@

# NAs in columns and rows with COG
table[,c(6,11)] <- NA
table[seq(4,16,4),] <- NA

printoutctab(table)

\begin{frame}[noframenumbering]
\frametitle{Outcomes}
\framesubtitle{Adulthood (BCS)}
\centering
\resizebox{\textwidth}{!}{
<<REGS_OUTC_BCS2_COG, results='asis', dependson="REGRESSIONS">>=

printoutctab(table)

@@

@
} %scalebox
\end{frame}
% 
% <<BCSEFA>>=
% @
% 
% <<BCSEFARES, dependson='BCSEFA'>>=
% @
% 
% <<MCSEFA>>=
% @
% 
% <<MCSEFARES, dependson='MCSEFA'>>=
% @
% 
% 
% \begin{frame}[noframenumbering]
% \frametitle{EFA}\label{numfac}
% \framesubtitle{Number of factors \hyperlink{efa}{\beamerreturnbutton{Back}}}
% \begin{center}
% <<EFATAB, dependson='MCSEFARES', results='asis'>>=
% # combine results
% efatab <- cbind(efatab.bcs,efatab.mcs)
% colnames(efatab) <- c("BCS All", "BCS M", "BCS F", "MCS All", "MCS M", "MCS F")
% 
% 
% print(xtable(efatab,
%                     align="L{.25\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
%                     digits = 0
%                     ),
%              size="\\small",
%              include.rownames=T,
%              include.colnames=T,
%              booktabs = T,
%              sanitize.text.function = function(x){x},
%              tabular.environment = "tabular", width = ".8\\textwidth",
%              floating = FALSE
% )
% @
% \end{center}
% \end{frame}

\begin{frame}
\frametitle{EFA}\label{meantab_fcc3}
\framesubtitle{Choice of common items (females) \hyperlink{meantab_fcc2}{\beamerreturnbutton{2 factors}}}
\centering
\resizebox{!}{4cm}{
<<MEANTAB_FCC3, dependson="MEANTABLE", results='asis'>>=
require(xtable)

table <- meantab_cc3[[2]] # females
table <- table[,!names(table) %in% c("cats", "fac", "bcs.num", "mcs.num")] # drop if unneeded

@@

\begin{frame}[noframenumbering]
\frametitle{MI}\label{miapp1}
\framesubtitle{Invariance steps \hyperlink{misteps}{\beamerreturnbutton{Back}}}
\Fontvi
\bi
\item \textbf{Configural} model
  \bi
  \item Minimally restrictive (all parameters free across cohorts)
  \item $\text{diag}(\bm{\Phi}) = \bm{I}, \quad \bm{\kappa}=\bm{0}, \quad \bm{\nu}=\bm{0}, \quad \text{and} \; \text{diag}(\bm{\Psi})=\bm{I}.$
  \ei
